\section{Discussion 1, Wednesday, January 22}

Set

\begin{itemize}
\item common set $\mathbf{Z}$, $\mathbf{Z}^{+}$, $\mathbf{N}$, $\mathbf{Q}$, $\mathbf{R}$, $\mathbf{R}^{+}$, $\mathbf{C}$
\item Operations $A\cap B$, $A\cup B$, $\overline A$, $A - B$
\end{itemize}

Cardinality 

$|A|$ number of distance elements

\begin{ex}
$|\{a, a, b, c\}| = 3$
\end{ex}

\begin{remark}
set order or duplication do not matter

$\{a, b, c\} = \{b, c, a\} = \{a, a, b, c\}$

\end{remark}

Ordered $n$-tuple

\begin{ex}
$(a, b, c) \neq (b, c, a)$

$(a, b, c) \neq (a, a, b, c)$
\end{ex}

\begin{prob}

Venn diagram

\end{prob}

\begin{remark}

Operators within the same pair of parentheses have the same level of priority; union and intersection have the same level of priority. Calculate from the innermost pair of parentheses.

\end{remark}


Power set $P(S)$

The set of all subsets of $S$

\begin{prob}
$|\O| = 0$

$|\{\O\}| = |\{\{\O\}\}| = 1$

$|\{\mathbf{Z}^{+}, 0, \mathbf{Z}^{-}\}| = 3$

$|\mathbf{Z}| = \infty$
\end{prob}

\begin{prob}

Show that if $A$, $B$, and $C$ are sets then ${\overline A \cap B \cap C} = {\overline A} \cup {\overline B} \cup {\overline C} \cup$

\end{prob}

Proof

Pick arbitrary element from left set, prove it is in the right set (subset). Do the same to the other side, proves they are equal.

Arbitrary $X$: could be anywhere $\cdots$ test every case

Functions

Must map to ONE element

Injective (one to one)

Each one mapped to unique element. Can have extra elements on right.

Surjective (onto)

Each element is mapped. No extra on the right.

Bijective (perfect match)

Injective and surjective

\begin{prob}

$f: \mathbf{Z} \rightarrow \mathbf{O}$ where $f(x) = 2x + 5$ is bijective

\end{prob}

\begin{prob}

$f: \mathbf{O} \rightarrow \mathbf{Z}^{+}$ where $f(x) = (x + 5)^2$ is neither

\end{prob}

Sequence

$\mathbf{N} \rightarrow S$

Fibonacci Sequence

$f_0 = 0$, $f_1 = 1$, $f_n = f_{n - 1} + f_{n - 2}$

\begin{prob}

$a_0 = 1$, $a_n = 2a_{n - 1}$, $a_n = 2^n$

\end{prob}

\begin{prob}

$a_0 = -7$, $a_{n + 1} = 10 - a_n$, $a_n = 5 - 12 (-1)^n$

\end{prob}


